---
title: "Replication of Study The Origins of the Shape Bias: Evidence From the Tsimane’ by Julian Jara-Ettinger et al. (2022 American Psychological Association)"
author: "Samah Abdelrahim (samahabd@stanford.edu)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
---

<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

##Introduction

The original paper is discussing the idea of the shape bias. They argue against the shape bias being a result of the language effect on thought, or an innate bias that facilitates the word meaning process. Instead, they hypothesize that shape bias emerges out of environmental factors that affect the kind of categories people encounter and talk about  in their daily life. Their hypothesis goes in line with the statistical regularities hypothesis which argues that the way people talk about things modulates this bias towards shape. 
Their work is a comparison between two populations that vary in the their level of industrialization which are people in the United States and people of the Tsimane’ who is less exposed to industrial artifacts. They compared the strength of shape bias between them. 
the original paper contains 7 experiments, but this replication project will only conduct the 5th experiment.

##Methods

###Power Analysis

Original effect size, power analysis for samples to achieve 80%, 90%, 95% power to detect that effect size.  Considerations of feasibility for selecting planned sample size.

###Planned Sample

The study was done with 144 english speaking adults from the US. They used a termination rule of having a sample size that is large enough to allow them to fully counterbalance all aspects of experimental design.

###Materials

For the replication experiment which is experiment 5 in the original article, the study was done online so they used photographs of solid objects that varied in shape, color, and material. 
I contacted the first author of the original paper to provide the photographs. 

###Procedure	

The experiment consisted of three example objects and three extension objects. The experiment was a one-shot learning trials and each participant completed one trial only. in this trial, each participant saw only one example object (counterbalanced across participants) and all extension objects. 
The participant sees a picture of the exempler object on a single screen with a sentence in the top that says something like "this is a yarn" or any other novel name in a sentence of the structure "this is a(n) x". Below, the text read “One of these is also a(n) x” along with three pictures: one shape-matched object, one material-matched object, and color-matched object. A text below also reads “Which one is the other x?” Participants were allowed to select one of the three objects.
They used three different possible labels, randomized across participants. The labels for the US participants were “koba,” “dax,” or “fep”. 

###Analysis Plan

Can also quote directly, though it is less often spelled out effectively for an analysis strategy section.  The key is to report an analysis strategy that is as close to the original - data cleaning rules, data exclusion rules, covariates, etc. - as possible.  

Data cleaning will be in the form of taking only a subset of the data frame that contains the features/columns of interest. No exclusion rule will be applied unless the participant didn't answer the Example at all. 
Then, I will present a simple descriptive statitics of the data by reporting the percentage of participants who extended by shape against those who extended by either color or material. 
Since i will only be testing English speaking adults from the US, there won't be any interaction effects of age or population. 
I will use a logistic model to test participant's preference for shape matches as a function of the exempler object. 

**Clarify key analysis of interest here**  You can also pre-specify additional analyses you plan to do.

###Differences from Original Study

No differences from the original paper are planned.


### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
  Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
 None


##Results


### Data preparation

Data preparation following the analysis plan.
	
```{r include=T}
###Data Preparation
#install.packages("lme4",repos="http://r-forge.r-project.org")
#install.packages("statmod")
#install.packages("lme4")

####Load Relevant Libraries and Functions
library(tidyverse)
library(foreign)
library(tidyr)
library(dplyr)
library(stringr) 
library(ggplot2)
library(here)
library(lme4)
####Import datare
d = read.csv("/Users/samahabdelrahim/Desktop/jara-ettinger2022/code/Jara-ettinger2022_November 27, 2022_14.55.csv")

head(d)

#### Data exclusion / filtering
d_filtered = d %>% select(c("ResponseId", "Q5", "Q9", "Q12"))
d_filtered_long = d_filtered %>% 
  pivot_longer(cols = -c("ResponseId") ,names_to = 'Example', values_to = 'Choice')

#use slice instead of filter

#### Prepare data for analysis - create columns etc.

d_filtered_long$Example[d_filtered_long$Example == "Q5"] <- "Yarn" #replace question number with the name of the object that is shown in that question

d_filtered_long$Example[d_filtered_long$Example == "Q9"] <- "Dax"
d_filtered_long$Example[d_filtered_long$Example == "Q12"] <- "Fep"

d_filtered_long$Choice[d_filtered_long$Choice == "IM_eY8Vwfih2OynNCm"] <- "Yarn"  #replace the serial number that identifies the answer choice with the name of the object that corresponds to that choice.

d_filtered_long$Choice[d_filtered_long$Choice == "IM_enDeJ2onL1JrLP8"] <- "Fep"
d_filtered_long$Choice[d_filtered_long$Choice == "IM_5A9xQWROP0v4aPA"] <- "Dax"

d_filtered_long = d_filtered_long %>% filter( !grepl("Import|Response",ResponseId)) #filter some unwanted rows out. 
#is.na(d_filtered_long$Choice)
d_filtered_long$Choice[d_filtered_long$Choice==""] <- NA #replace empty slots with NA
d_filtered_long = d_filtered_long %>% filter(!is.na(Choice)) # maintain only the questions that have answers, i.e. the question that appeared to each participant in the random trial. 

d_filtered_long = d_filtered_long %>% 
  mutate(Match = ifelse(Example == Choice, "Shape", 
                        ifelse(Example == "Yarn" & Choice == "Fep","Color", 
                               ifelse(Example == "Yarn" & Choice =="Dax", "Material", 
                                      ifelse(Example =="Dax" & Choice == "yarn", "Color",
                                             ifelse(Example =="Dax" & Choice == "Fep", "Material",
                                                    ifelse(Example == "Fep" & Choice =="Dax", "Color", "Material"
                                                           )
                                                    )
                                             )
         )
         )
         )
  )
         

d_filtered_long = d_filtered_long %>%
  mutate(Location = "US", Experiment= "US_Adults_Objects") %>% 
  select(c("Experiment", "Example", "Match","Location" ))
colnames(d_filtered_long)[3] = "Choice"

d_filtered_long_summarized = d_filtered_long %>% 
  group_by(Choice, Location) %>%
  summarise( n = n(), Percentage = n / nrow(d_filtered_long))



# length() only gets distinct un-redundant

head(d_filtered_long)

```
``` {r include = T}
# importing the Tsimane data from original paper

d_Tsimane = read.csv("/Users/samahabdelrahim/Desktop/jara-ettinger2022/code/ShapeBias_Data_original.csv")
d_Tsimane_filtered = d_Tsimane %>% select(c("Experiment", "Example", "Choice","Location"))

d_Tsimane_summarized = d_Tsimane_filtered %>% 
  group_by(Choice, Location) %>%
  summarise( n = n(), Percentage = n / nrow(d_Tsimane_filtered))

df_combined <- rbind(d_filtered_long, d_Tsimane_filtered)
df_combined_summarised <- rbind(d_filtered_long_summarized, d_Tsimane_summarized)

df_combined$Example[df_combined$Example == "Yarn"] <- "Arch" #replace question number with the name of the object that is shown in that question

df_combined$Example[df_combined$Example == "Dax"] <- "Snowman"
df_combined$Example[df_combined$Example == "Fep"] <- "Lamp"

```
### Confirmatory analysis

The analyses as specified in the analysis plan.  
If participants used shape matches to extend the word more than they used material or color, then the result is replicated. No comparison will be shown here since i am only testing one population (US adults).

``` {r include = T}
#d_filtered_long_summarized = df_combined %>%
 # group_by(Example, Choice, Location) %>%
  #summarise(n = n(), avg = n / count(Location == Location))
  #summarise(counter = n(Match == "Shape"))

#using only us data
#graph1 <- ggplot(d_filtered_long, aes(x=Choice, fill=Example)) +
 # geom_bar(position="dodge")  
#graph1


ggplot(df_combined, aes(x=Example, fill = Choice)) +
  geom_bar(position = "dodge", width = .7) +
  theme_minimal() +
  scale_fill_manual(values = c("#B4DCB9", "#7F92B8", "#6C6969")) +
  scale_x_discrete("Example") +
  scale_y_continuous("Frequency of choices out of - ") +
  facet_wrap(~Location) +
  labs(title = "Word - Object categorization", subtitle = "Shape bias")


ggplot(df_combined_summarised, aes(x = Location, fill = Choice)) +
         geom_bar( aes(y = Percentage), stat = "identity", width = 0.7) +
         theme_minimal() +
         scale_fill_manual(values = c("#B4DCB9", "#7F92B8", "#6C6969" ))
         
```

``` {r include = T}
# Statistical analysis

data_new <- d_filtered_long      
data_new$Choice[data_new$Choice == "Shape"] <- 1
data_new$Choice[data_new$Choice == "Material"] <- 0

# Duplicate data
#data_new[is.na(data_new) | data_new == "Inf"] <- NA
#model2 = glmer (Choice ~ Question, data_new)
#model1 = lm(Choice ~ Example, data_new)
#model1
#summary(model1)
#chisq.test(d_filtered_long$Choice, d_filtered_long$Example, correct=FALSE)

#using the combined data
m <- glmer(Choice=="Shape" ~  Location  + (1 + Location| Example) + (1 | Experiment),df_combined,family="binomial")
summary(m)
```
*Side-by-side graph with original graph is ideal here*
 
 Matches by shape are considerbaly more than matches by other attributes like material or color. 
 
###Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.
